---
title: "Bases para la inferencia estadística: Muestreando distribuciones"
runtime: shiny
output:
  html_document:
    css: www/lab.css
    highlight: pygments
    theme: cerulean
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
library(dplyr)
library(ggplot2)
library(mosaic)
load("ames.RData")
```

<div id="template">
**Tu reporte de laboratorio reproducible:** Antes de empezar descarga la 
plantilla para este laboratorio. Recuerda que todo el código y tus respuestas
van en este documento:

```{r load-template, eval = FALSE}
download.file("https://git.io/lab3", destfile = "lab3.Rmd")
```
</div>

En este laboratorio vamos a investigar las maneras en los cuales los estadísticos
de una muestra aleatoria de datos puede servir como estimadores puntales de 
los parametros poblacionales. Estamos interesados en formular una *distribución
muestral* de nuestro estimador para conocer las propiedades del estimado, tales
como su distribución.

<div id="boxedtext">
**Configurando la semilla:** Vamos a tomar algunas muestras aleatorios y crear
distribuciones muestrales en este laboratorio, lo que significa que deberías 
configurar la semilla al principio del laboratorio. Si este concepto es nuevo,
revisa el laboratorio anterior y pregunta a los docentes.
</div>

## Los datos

Vamos a considerar los datos inmobiliarios de la ciudad de Ames, Iowa. Los
detalles de cada transacción inmobiliara de Ames es registrada por la oficina
del Asesor de la Ciudad. En particular, nuestro foco para este laboratorio será
todas las ventas de hogares residenciales en Ames entre 2006 y 2010. Esta
colección representa nuestra población de interes. En este laboratorio nos 
gustaría apredner sobre esta ventas de hogares tomando muestras más pequeñas de
la población total. Carguemos los datos.



```{r load-data}
load(url("https://git.io/ames.RData"))
```

Vemos que hay varias variables en el conjunto de datos, suficientes como para
hacer un análisis en profundidad. Para este laboratorio, nos vamos a restringir
solo a dos variables: el area total sobre el suelo de la casa en pies cuadrados
(`area`) y el precio de venta (`price`).

Podemos explorar la distribución de las areas de las casas en la población de
ventas de casa visualmente y con estadísticos de resumen. Primero creemos una
visualización, un histrograma:

```{r area-hist}
ggplot(data = ames, aes(x = area)) +
  geom_histogram(binwidth = 250)
```

Vamos a obtener algunas medidas de resumen. Noten que podemos hacer esto 
usando la función `summarise`. Podemos calcular tantas medidas como queramos
usando esta función, y solo encandenando los resultados. Algunas de las 
funciones abajo se explican solas (`mean`, `median`, `sd`, `min`, `max`).
`IQR` es el rango intercuartiles. Una función nueva que encontramos aquí es 
`quantile`, la cual nos permite calculara los valores correspondientes a 
cuantiles de la distribución. Por ejemplo, `quantile(x, 0.25)` dará el valor
de corte para el percentil 25° (Q1) de la distribución de x. Encontrar estos
valores es útil para describir la distribución, ya que podemos usarlos para 
descripciones como *"la mitad de las casas tienen areas entre tanto y tanto 
pies cuadrados"*


```{r area-stats}
ames %>%
  summarise(mu = mean(area), pop_med = median(area), 
            sigma = sd(area), pop_iqr = IQR(area),
            pop_min = min(area), pop_max = max(area),
            pop_q1 = quantile(area, 0.25),  # primer cuartil, percentil 25°
            pop_q3 = quantile(area, 0.75))  # tercer cuartil, percentil 75°
```

1.  Describe la distribución de la población usando la visualización y las
    medidas de resumen. No hace falta usar todas las medidas de resumen en la
    descripción, deberán decidir cuales son relevantes basadas en la forma de 
    la distribución. Asegurate de incluir el gráfico y las medidas de resumen
    en su informe además de la descripción.


## La distribución de muestreo desconocida

En este laboratorio tenemos acceso a toda la población, pero esto raramente
sucede en la vida real. Conseguir la información de toda una población es
extremadamente costoso o imposible. Debido a esto, normalmente tomamos una 
muestra de la población y la usamos para entender las propiedades de la
población.

Si estuviesemos interesados en estimar la media del area en Ames basados en una
muestra, podemos usar el siguiente commando para muestrear la población.

```{r samp1}
samp1 <- ames %>%
  sample_n(50)
```

This command collects a simple random sample of size 50 from the `ames` dataset 
`area`, which is assigned to `samp1`. This is like going into the City 
Assessor's database and pulling up the files on 50 random home sales. Working 
with these 50 files would be considerably simpler than working with all 2930 
home sales.

1.  Describe the distribution of area in this sample. How does it compare to the 
    distribution of the population? **Hint:** `sample_n` function takes a random
    sample of observations (i.e. rows) from the dataset, you can still refer to 
    the variables in the dataset with the same names. Code you used in the 
    previous exercise will also be helpful for visualizing and summarizing the sample,
    however be careful to not label values `mu` and `sigma` anymore since these
    are sample statistics, not population parameters. You can customize the labels
    of any of the statistics to indicate that these come from the sample.

If we're interested in estimating the average living area in homes in Ames 
using the sample, our best single guess is the sample mean.

```{r mean-samp1}
samp1 %>%
  summarise(x_bar = mean(area))
```

Depending on which 50 homes you selected, your estimate could be a bit above 
or a bit below the true population mean of 1499.69 square feet. In general, 
though, the sample mean turns out to be a pretty good estimate of the average 
living area, and we were able to get it by sampling less than 3\% of the 
population.

1.  Would you expect the mean of your sample to match the mean of another team's 
    sample? Why, or why not? If the answer is no, would you expect the means to
    just be somewhat different or very different? Ask a neighboring team to confirm 
    your answer.

1.  Take a second sample, also of size 50, and call it `samp2`. How does the 
    mean of `samp2` compare with the mean of `samp1`? Suppose we took two 
    more samples, one of size 100 and one of size 1000. Which would you think 
    would provide a more accurate estimate of the population mean?

Not surprisingly, every time we take another random sample, we get a different 
sample mean. It's useful to get a sense of just how much variability we 
should expect when estimating the population mean this way. The distribution 
of sample means, called the *sampling distribution*, can help us understand 
this variability. In this lab, because we have access to the population, we 
can build up the sampling distribution for the sample mean by repeating the 
above steps many times. Here we will generate 2000 samples and compute the 
sample mean of each. Note that we are sampling with replacement, 
`replace = TRUE` since sampling distributions are constructed with sampling
with replacement.

```{r loop}
sample_means50 <- do(2000) * ames %>% 
                              sample_n(50, replace = TRUE) %>%
                              summarise(x_bar = mean(area))

qplot(data = sample_means50, x = x_bar)
```

Here we use R to take 2000 samples of size 50 from the population, calculate 
the mean of each sample, and store each result in a vector called 
`sample_means50`. On the next page, we'll review how this set of code works.

1.  How many elements are there in `sample_means50`?  Describe the sampling 
    distribution, and be sure to specifically note its center. Make sure to include 
    a plot of the distribution in your answer.

## Interlude: The `do` function

Let's take a break from the statistics for a moment to let that last block 
of code sink in. The idea behind the do function is *repetition*: it allows 
you to execute a line of code as many times as you want and put the results 
in a data frame.  In the case above, we wanted to repeatedly take a random 
sample of size 50 from `area` and then save the mean of that sample into 
the `sample_means50` vector.

Without the `do` function, this would be painful. First, we'd have to create 
an empty vector filled with 0s to hold the 2000 sample means. Then, we'd have 
to compute each of the 2000 sample means one line at a time, putting them 
individually into the slots of the `sample_means50` vector:

```{r do-long}
sample_means50 <- rep(NA, 2000)

sample_means50[1] <-  ames %>%
                        sample_n(50, replace = TRUE) %>%
                        summarise(x_bar = mean(area))
sample_means50[2] <- ames %>%
                        sample_n(50, replace = TRUE) %>%
                        summarise(x_bar = mean(area))
sample_means50[3] <- ames %>%
                        sample_n(50, replace = TRUE) %>%
                        summarise(x_bar = mean(area))
sample_means50[4] <- ames %>%
                        sample_n(50, replace = TRUE) %>%
                        summarise(x_bar = mean(area))
# ...and so on, 2000 times
```

With the `do` function, these thousands of lines of code are compressed 
into one line:

```{r do-again}
sample_means50 <- do(2000) * ames %>%
                              sample_n(50, replace = TRUE) %>%
                              summarise(x_bar = mean(area))
```

Note that for each of the 2.000 times we computed a mean, we did so from a 
**different** sample!

1.  To make sure you understand how sampling distributions are built, and exactly
    what the `sample_n` and `do` function do, try modifying the code to create a
    sampling distribution of **25 sample means** from **samples of size 10**, 
    and put them in a data frame named `sample_means_small`. Print the output. 
    How many observations are there in this object called `sample_means_small`? 
    What does each observation represent?

## Sample size and the sampling distribution

Mechanics aside, let's return to the reason we used the `do` function: to compute a 
sampling distribution, specifically, this one.

```{r hist}
qplot(data = sample_means50, x = x_bar, geom = "histogram")
```

The sampling distribution that we computed tells us much about estimating 
the average living area in homes in Ames.  Because the sample mean is an 
unbiased estimator, the sampling distribution is centered at the true average 
living area of the population, and the spread of the distribution 
indicates how much variability is induced by sampling only 50 home sales.

In the remainder of this section we will work on getting a sense of the effect that 
sample size has on our sampling distribution.

1.  Use the app below to create sampling distributions of means of `area`s from
    samples of size 10, 50, and 100. Use 5,000 simulations. What does each 
    observation in the sampling distribution represent? How does the mean, standard
    error, and shape of the sampling distribution change as the sample size
    increases? How (if at all) do these values change if you increase the number 
    of simulations? (You do not need to include plots in your answer.)
    
```{r shiny, echo=FALSE, eval=TRUE}
shinyApp(
  ui <- fluidPage(
    
    # Sidebar with a slider input for number of bins 
    sidebarLayout(
      sidebarPanel(
        
        selectInput("selected_var",
                    "Variable:",
                    choices = list("area", "price"),
                    selected = "area"),         
        
        numericInput("n_samp",
                     "Sample size:",
                     min = 1,
                     max = nrow(ames),
                     value = 30),
        
        numericInput("n_sim",
                     "Number of samples:",
                     min = 1,
                     max = 15000,
                     value = 5000) 
        
      ),
      
      # Show a plot of the generated distribution
      mainPanel(
        plotOutput("sampling_plot"),
        verbatimTextOutput("sampling_mean"),
        verbatimTextOutput("sampling_se")
      )
    )
  ),
  
  # Define server logic required to draw a histogram
  server <- function(input, output) {
    
    # create sampling distribution
    sampling_dist <- reactive({
      ames[[input$selected_var]] %>%
        sample(size = input$n_samp * input$n_sim, replace = TRUE) %>%
        matrix(ncol = input$n_samp) %>%
        rowMeans() %>%
        data.frame(x_bar = .)
    })
    
    # plot sampling distribution
    output$sampling_plot <- renderPlot({
      x_min <- quantile(ames[[input$selected_var]], 0.1)
      x_max <- quantile(ames[[input$selected_var]], 0.9)
      
      ggplot(sampling_dist(), aes(x = x_bar)) +
        geom_histogram() +
        xlim(x_min, x_max) +
        ylim(0, input$n_sim * 0.35) +
        ggtitle(paste0("Sampling distribution of mean ", 
                       input$selected_var, " (n = ", input$n_samp, ")")) +
        xlab(paste("mean", input$selected_var)) +
        theme(plot.title = element_text(face = "bold", size = 16))
    })
    
    # mean of sampling distribution
    output$sampling_mean <- renderText({
      paste0("mean of sampling distribution = ", round(mean(sampling_dist()$x_bar), 2))
    })
    
    # mean of sampling distribution
    output$sampling_se <- renderText({
      paste0("SE of sampling distribution = ", round(sd(sampling_dist()$x_bar), 2))
    })
  },
  
  options = list(height = 500) 
)
```


* * *

<div class="oyo">

## On your own

So far, we have only focused on estimating the mean living area in homes in 
Ames. Now you'll try to estimate the mean home price.

Note that while you might be able to answer some of these questions using the app
you are expected to write the required code and produce the necessary plots and
summary statistics. You are welcomed to use the app for exploration.

-   Take a sample of size 15 from the population and calculate the mean `price`
    of the homes in this sample. Using this sample, what is your best point estimate
    of the population mean of prices of homes?

-   Since you have access to the population, simulate the sampling 
    distribution for $\bar{x}_{price}$ by taking 2000 samples from the 
    population of size 15 and computing 2000 sample means. Store these means 
    in a vector called `sample_means15`. Plot the data, then describe the 
    shape of this sampling distribution. Based on this sampling distribution, 
    what would you guess the mean home price of the population to be? Finally, 
    calculate and report the population mean.

-   Change your sample size from 15 to 150, then compute the sampling 
    distribution using the same method as above, and store these means in a 
    new vector called `sample_means150`. Describe the shape of this sampling 
    distribution, and compare it to the sampling distribution for a sample 
    size of 15. Based on this sampling distribution, what would you guess to 
    be the mean sale price of homes in Ames?

-   Of the sampling distributions from 2 and 3, which has a smaller spread?  If
    we're concerned with making estimates that are more often close to the 
    true value, would we prefer a sampling distribution with a large or small spread?

</div>

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0).
This lab was written for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel.
</div>